---
title: "Covid 19 case prediction (linear regression models and eveluation)"
author: "Gloria Xiang"
date: "2022-12-03"
output:
  pdf_document: default
---

The covid19 dataset comes from real covid 19 statistics among different countries.\
write.csv(covid_predict, "/Users/miao/Desktop/covid_predict.csv", row.names=FALSE)


```{r include=FALSE}
#set country and time frame (only after 10/01/2021 had completed records)
library(COVID19)
covid<-covid19("US",start="2021-10-01", end="2022-03-31")
colnames(covid)
```

## Introduction: Descriptive analysis (linear plot)
```{r}
plot(covid$deaths~covid$date,type="l", xlab="Month")
plot(covid$deaths~covid$tests,type="l")
plot(covid$deaths~covid$confirmed,type="l")
```

Remove NA/character variables.\

```{r}
covid2<-covid[,-c(1,5,12:19,21:26,30:47) ]
covid2$date<-as.numeric(covid2$date)
```

## Estimate:
Create a model with deaths as the outcome variable and the rest of variables as predictors. We do the log transfer for test and vaccine variables in the full model.\

```{r}
lmod_full<-lm(sqrt(deaths)~log(tests)+log(vaccines)+confirmed+people_vaccinated+
           people_fully_vaccinated+hosp+icu+international_movement_restrictions+
           government_response_index+stringency_index+containment_health_index,
         data=covid2)
summary(lmod_full)
```


## Model Selection:

Use BIC values to select variables.\

```{r echo=TRUE}
library(leaps)
b <- regsubsets(deaths~log(tests)+log(vaccines)+confirmed+people_vaccinated+
           people_fully_vaccinated+hosp+icu+international_movement_restrictions+
           government_response_index+stringency_index+containment_health_index,
         data=covid2)
rs <- summary(b)
# 1 subsets of each size up to 8
BIC <- 182*log(rs$rss/182) + (2:9)*log(182)
plot(BIC ~I(1:8), ylab="BIC", xlab="Number of Predictors")
```
The  BIC selection shows the more predictors, the lower BIC

Use AIC forward selection to select variables.\

```{r echo=TRUE}
fit_null <- lm(deaths ~ 1, data = covid2)
fit_AIC <- step(fit_null, 
                scope = list(lower = fit_null,
                             upper = lmod_full),
                direction = "forward",criterion = "AIC")
summary(fit_AIC)
```

Multiple R-squared is 0.9999. This is an over-fitted model, so we need to do the diagnostics and remove some predictors.

## Diagnostics
Variance constant
```{r}
plot(fit_AIC$fitted.values, fit_AIC$residuals,ylab="Residual", xlab="Fitted") 
```
There is no heteroskedasticity. 

Normality test
```{r}
qqnorm(residuals(fit_AIC),ylab="Residuals",main="")
qqline(residuals(fit_AIC))
```
Q-Q plot shows the residuals are normally distributed. 

Shapiro-Wilk test
```{r}
shapiro.test(residuals(fit_AIC))
```
Do not reject null hypothesis. The residuals are normally distributed.\

VIF check collinearity
```{r}
library(faraway)  
x <- model.matrix(fit_AIC)[,-1] 
vif(x)
```

The result of VIFs shows multicollinearity (big VIFs). Next, remove variables with big VIF, then do inference and prediction.  

## model 1
## using the model from AIC selection
## Inference:
95% Confidence interval 
```{r}
confint(fit_AIC)
```

## Prediction:
Using real data from "2022-04-01" to "2022-06-18", because there is no "test" data available after 2022-06-18. 

```{r}
x0 <-covid19("US",start="2022-04-01", end="2022-06-18")[,-c(1,5,12:19,21:26,30:47) ]
p<-predict(fit_AIC,new=x0)
```

```{r}
plot(x0$deaths~x0$date,type="l", xlab="Month")
i <- order(x0$date)
lines(x0$date[i],p[i],lty=2)
legend("topleft", inset=.05, title="Deaths of Covid19",
c("Real","Predicted"), lty=1:2, horiz=TRUE)
```

Predicted values are smaller than real data.

## model 2
## Exclude predictors with big VIF
```{r}
VIF_lomd<-lm(deaths~confirmed+people_fully_vaccinated+hosp
             +government_response_index+stringency_index,data=covid2)

summary(VIF_lomd)
```

95% Confidence interval 
```{r}
confint(VIF_lomd)
```
Using real data from "2022-04-01" to "2022-09-30"
```{r}
covid_predict <-covid19("US",start="2022-04-01", end="2022-09-30")[,-c(1,5,12:19,21:26,30:47) ]
p<-predict(VIF_lomd,new=covid_predict)
```

```{r}
plot(covid_predict$deaths~covid_predict$date,type="l", xlab="Month")
i <- order(covid_predict$date)
lines(covid_predict$date[i],p[i],lty=2)
legend("topleft", inset=.05, title="Deaths of Covid19",
c("Real","Predicted"), lty=1:2, horiz=TRUE)
```
Similar results to model 1 (fit_AIC)


## Transformation
## Fianl model 1
Convert all variables to standard units
```{r}
scsav <- data.frame(scale(covid2))
Final_1<-lm(deaths~confirmed+people_fully_vaccinated+hosp+
              government_response_index+stringency_index,data=scsav)
summary(Final_1)
```

VIF check collinearity
```{r}
library(faraway)  
x <- model.matrix(Final_1)[,-1] 
vif(x)
```

## Prediction:
Using real data from "2022-04-01" to "2022-06-18", because there is no "test" data available after 2022-06-18.(international_movement_restrictions unchanged during this period, so the prediction excluded this variable)

Using real data from "2022-04-01" to "2022-09-30"

```{r}
covid_predict <-covid19("US",start="2022-04-01",
                        end="2022-09-30")[,-c(1,5,12:19,21:26,30:47) ]

covid_predict[,2:13]<-scale(covid_predict[,2:13])
p<-predict(Final_1,new=covid_predict)
```

## inference-1
```{r}
confint(Final_1)
```

```{r}
plot(covid_predict$deaths~covid_predict$date,type="l", xlab="Month")
i <- order(covid_predict$date)
lines(covid_predict$date[i],p[i],lty=2)
legend("topleft", inset=.05, title="Deaths of Covid19",
c("Real","Predicted1"), lty=1:2, horiz=TRUE)
```


Transformation2: box-cox
```{r}
Final_2 <- lm(sqrt(deaths)~confirmed+people_fully_vaccinated+hosp+ government_response_index+stringency_index,data=covid2)
summary(Final_2)
```

```{r}
x0 <-covid19("US",start="2022-04-01", end="2022-06-18")[,-c(1,5,12:19,21:26,30:47) ]
x0$deaths<-sqrt(x0$deaths)
p<-predict(lmod_full,new=x0)
```

```{r}
plot(x0$deaths~x0$date,type="l", xlab="Month")
i <- order(x0$date)
lines(x0$date[i],p[i],lty=2)
legend("topleft", inset=.05, title="Deaths of Covid19",
c("Real","Predicted2"), lty=1:2, horiz=TRUE)
```

```{r}
# Box-Cox 
library(MASS)
box<-boxcox(lmod_full, plotit=T,lambda=seq(0,1,by=0.1))

cox <- data.frame(box$x, box$y)
cox2 = cox[with(bc,order(-cox$box.y)),]
lambda = cox2[1,"box.x"]
covid2$deaths <- (covid2$deaths^lambda-1)/lambda
logtrans(lmod_full,plotit=TRUE, alpha=seq(-min(covid2$deaths)+0.001,0,by=0.01))
```

Interpretation of the final model:
The model has an R square around 0.99 in both predictions, which means, unfortunately, the models have over-fitting problem. It could because of: 
  - 1. The death number kept a stable increasing status. The research time is not at the very beginning of the outbreak, vibration in cases is much more seldom; 
  - 2. The dataset is a primary dataset, it hardly has any other evaluations, simply using a linear regression model could be hard to predict precisely.

Future analysis suggestions: 
  - From Study design, the model should have better performance by selecting train and test data randomly, since it could decrease the over-fitting problem by randomly choosing the date;
  - Use the death rate(death/confirmation) instead of deaths number, and calculate the new-confirmed (confirmed - confirmed from yesterday), new-deaths (deaths- deaths from yesterday) in future study.
  - Conduct Time Series Analysis to analysis the relationship between time pattern & deaths.
